---
title: "PD bulk CUT&Tag preprocessing and quality control"
author: "Samantha Lee"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Read data in 

Load required libraries
```{r libs}
library(here) # dir calling
library(tidyverse) # data wrangling 
library(plotly) # interactive plotting
library(readxl) # read xls files
library(DESeq2) # depth normalization + diff analysis
library(pvca) # examine variance in pcs
source(here("scripts","r","heat_scree_plot_corr.r"))
```

Read in participant metadata
```{r data_in_meta}
meta <- read.csv(here("resources",
                      "metadata",
                      "metadata_participants.csv")) %>%
  mutate(group = ifelse(grepl("PDC", pid), "control", "pd"))
```

Read in fastQC data for merged fastq files. 
```{r data_in_merged_fastqc}
merged_fastqc <- read_table(here("data_out",
                                 "020_merged",
                                 "021_fastqc",
                                 "multiqc_data",
                                 "multiqc_fastqc.txt"), 
                            col_names = T)
```

Read in fastQC data for trimmed fastq files. 
```{r data_in_trimmed_fastqc}
trimmed_fastqc <- read_table(here("data_out",
                                  "030_trimmed",
                                  "031_fastqc",
                                  "multiqc_data",
                                  "multiqc_fastqc.txt"), 
                             col_names = T)  
```

Read in samtools stat data for aligned reads. 
```{r data_in_aligned_samtools_stats}
collect_sn_metrics <- function(
    indir,
    pattern,   # files ending with .stats
    recursive = TRUE,
    out_csv = "sn_metrics.csv"
) {
  files <- list.files(indir, pattern = pattern, recursive = recursive, full.names = TRUE)
  if (length(files) == 0L) stop("No files matched '", pattern, "' in ", indir)
  
  parse_one <- function(path) {
    lns <- readLines(path, warn = FALSE)
    sn <- grep("^SN\\t", lns, value = TRUE)
    if (!length(sn)) return(NULL)
    
    # Clean: drop leading "SN\t", remove inline comments after a tab (e.g., "\t# ...")
    sn <- sub("^SN\\t", "", sn)
    sn <- sub("\\t#.*$", "", sn)
    
    parts <- strsplit(sn, "\t", fixed = TRUE)
    metrics <- vapply(parts, function(x) sub(":$", "", x[1]), character(1))
    values  <- vapply(parts, function(x) x[2], character(1))
    
    # Best-effort numeric casting (handles integers and scientific notation)
    values_tc <- type.convert(values, as.is = TRUE)
    
    # Build one-row data.frame
    df <- as.data.frame(as.list(values_tc), stringsAsFactors = FALSE, optional = TRUE)
    names(df) <- metrics
    
    # Sample name: prefer IGF####### in filename; else stem before '_' ; else stem
    fname <- basename(path)
    igf <- regmatches(fname, regexpr("IGF\\d+", fname))
    sample <- if (length(igf) && igf != "") {
      igf
    } else {
      stem <- sub("\\.[^.]*$", "", fname)
      if (grepl("_", stem)) sub("_.*$", "", stem) else stem
    }
    
    df$sample <- sample
    df
  }
  
  dfs <- Filter(Negate(is.null), lapply(files, parse_one))
  if (!length(dfs)) stop("Parsed no SN metrics from matched files.")
  
  # Row-bind (fill missing columns), move 'sample' first, sort other columns
  all_cols <- unique(c("sample", sort(unique(unlist(lapply(dfs, names)))) ))
  dfs2 <- lapply(dfs, function(d) {
    missing <- setdiff(all_cols, names(d))
    if (length(missing)) d[missing] <- NA
    d[all_cols]
  })
  
  # write out
  out <- unique(do.call(rbind, dfs2))
  # utils::write.csv(out, out_csv, row.names = FALSE)
  # message("Wrote ", out_csv, " with ", nrow(out), " samples and ", ncol(out)-1, " metrics.")
  # invisible(out)
}

# get file stats
aligned_stats <- collect_sn_metrics(
  indir   = here("data_out",
                 "040_aligned",
                 "041_samtools_stats"),
  pattern = "\\.txt$",        
  recursive = TRUE
)
```

Read in fastQC data for aligned and deduplicated reads. 
```{r data_in_deduplicated_fastqc}
deduplicated_fastqc <- read_table(here("data_out",
                                       "050_deduplicated",
                                       "051_fastqc",
                                       "multiqc_data",
                                       "multiqc_fastqc.txt"), 
                                  col_names = T)    
```

Read in samtools stats data for aligned and deduplicated reads. 
```{r data_in_deduplicated_samtools_stats}
deduplicated_stats <- collect_sn_metrics(
  indir   = here("data_out",
                 "050_deduplicated",
                 "053_samtools_stats"),
  pattern = "\\.txt$",        
  recursive = TRUE
)  
```

Read in samtools stats data for aligned, deduplicated, and quality filtered reads.
```{r data_in_filtered_samtools_stats}
filtered_stats <- collect_sn_metrics(
  indir   = here("data_out",
                 "060_filtered",
                 "061_samtools_stats"),
  pattern = "\\.txt$",        
  recursive = TRUE
)  
```

Read in fragment length data (TLEN from quality filtered sam files).
```{r data_in_filtered_fragment_lengths}
sample_from_file <- function(path) {
  fname <- basename(path)
  igf <- regmatches(fname, regexpr("IGF\\d+", fname))
  if (length(igf) && igf != "") return(igf)
  stem <- sub("\\.[^.]*$", "", fname)
  if (grepl("_", stem)) sub("_.*$", "", stem) else stem
}

# Read one file: 2 columns (length, count), no header
read_frag_file <- function(path) {
  df <- read.table(path, header = FALSE,
                   col.names = c("insert_size", "count"),
                   colClasses = c("integer", "numeric"))
  df$sample <- sample_from_file(path)
  df
}

# Collect all files that match a regex (change pattern as needed)
collect_fragments <- function(indir,
                              pattern = "\\.txt$",   # <-- change to your extension
                              recursive = TRUE) {
  files <- list.files(indir, pattern = pattern, recursive = recursive, full.names = TRUE)
  if (!length(files)) stop("No files matched regex pattern: ", pattern, " in ", indir)
  
  dfs <- lapply(files, read_frag_file)
  do.call(rbind, dfs)
}

# get frag lengths
frag_lengths <- collect_fragments(here("data_out",
                                       "060_filtered",
                                       "062_fragment_lengths"), 
                                  pattern = "\\.txt$")  
```

Read in FRiP scores. 
```{r data_in_peaks_frip_scores}
frip_in <- function(f) {
  read_table(f,
             skip = 0, 
             col_names = T)
}

frip_scores <- map_dfr(list.files(here("data_out",
                                       "070_peaks", 
                                       "072_frip_scores"),
                                  pattern = "^.*\\.txt$",
                                  full.names = TRUE), 
                       frip_in) 
```

## Total read count (forward and reverse) across preprocessing steps

```{r plot_read_counts}
read_count_plot <- data.frame(sample = filtered_stats$sample,
                              merged_reads = merged_fastqc %>% subset(grepl("R1", Sample)) %>% pull(Sequences_1)*2,
                              trimmed_reads = trimmed_fastqc %>% subset(grepl("R1", Sample)) %>% pull(Sequences_1)*2,
                              aligned_reads = aligned_stats$`reads mapped and paired`,
                              deduplicated_reads = deduplicated_stats$`reads mapped and paired`,
                              filtered_reads = filtered_stats$`reads mapped and paired`) %>%
  pivot_longer(-sample) %>%
  mutate(name = factor(name, levels = c("merged_reads", 
                                        "trimmed_reads",
                                        "aligned_reads",
                                        "deduplicated_reads",
                                        "filtered_reads"))) %>%
  ggplot(aes(x = name, 
             y= value/1000000,
             fill = name)) +
  geom_col() +
  geom_hline(yintercept = 10, 
             color = "#e1235c", 
             linetype = "dashed", 
             linewidth = 0.25) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = -90,
                                   hjust = 0,
                                   vjust = 0.5)) +
  facet_wrap(~sample) +
  scale_x_discrete(name = "Preprocessing step",
                   labels = c("Raw", "Trimmed", "Aligned", "Deduplicated", "Filtered")) +
  scale_y_continuous(name = "Total reads (millions)") +
  scale_fill_manual(values = c("#00033f",
                               "#002965",
                               "#1d4f8b",
                               "#4375b1",
                               "#689ad6"))
ggplotly(read_count_plot)
```

**Interpretation**: Read counts look good. All samples have >10 million reads after alignment, deduplicate, and quality filtering. All samples pass. 


## Distribution of fragment lengths

```{r plot_fragment_lengths}
tlen_plot <- ggplot(frag_lengths,
                    aes(x=insert_size,
                        y = count,
                        color=sample)) +
  geom_line() +
  # geom_smooth(se = FALSE, method = "loess", span = 0.15) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c(RColorBrewer::brewer.pal(3, "Greens"),
                                RColorBrewer::brewer.pal(3, "Oranges"),
                                RColorBrewer::brewer.pal(3, "Blues"),
                                RColorBrewer::brewer.pal(3, "Purples"),
                                RColorBrewer::brewer.pal(3, "Reds"),
                                RColorBrewer::brewer.pal(3, "Greens"),
                                RColorBrewer::brewer.pal(3, "Oranges"),
                                RColorBrewer::brewer.pal(3, "Blues"),
                                RColorBrewer::brewer.pal(3, "Purples"),
                                RColorBrewer::brewer.pal(3, "Reds"))) +
  scale_y_continuous(name = "Count") +
  scale_x_continuous(name = "Fragment size (bp)")

ggplotly(tlen_plot)
```


**Interpretation**: Most distributions look reasonable. A few are skewed towards smaller fragments. All samples pass based on visual checks. 

## Fraction of Reads in Peaks (FRiP) scores

```{r frip_plot}
frip_plot <- ggplot(frip_scores,
                    aes(x = file, 
                        y = percent)) +
  geom_col(fill = "#002965") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, 
                                   hjust = 0,
                                   vjust = 0.5)) +
  scale_x_discrete(name = "Sample") +
  scale_y_continuous("Fraction of reads in peaks(%)") +
  geom_hline(yintercept = 10, 
             color = "#e1235c", 
             linetype = "dashed", 
             linewidth = 0.5) +
  geom_hline(yintercept = 20, 
             color = "#e1235c", 
             linetype = "dotted", 
             linewidth = 0.5) 

ggplotly(frip_plot)
```


**Interpretation**: FRiP >20% considered good and >10% acceptable. Samples ending in 15, 16, and 17 belong to the same control participant and all exhibit FRiP <10%. Discard IGF136815, IGF136816, and IGF136817 from all downstream analyses.  

## Consesus features across all samples

Clustering is based on sample feature (peak) counts across all consensus peaks (identified using all samples except for IGF136815, IGF136816, and IGF136817).

Read in feature count data for all sampla
```{r data_in_peaks_feature_counts_all}
# list files and derive sample names
tsv_files <- list.files(here("data_out",
                             "070_peaks",
                             "077_peaks_feature_counts_all"), 
                        pattern = "*.tsv$", 
                        full.names = TRUE)

# get sample names from file names
sample_names <- sub("_feature_counts_all_multicov\\.tsv$", 
                    "", 
                    basename(tsv_files))

# read each file, take the last column, and column-bind into one data frame
features_all <- map_dfc(
  tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = F) %>% select(dplyr::last_col())) 

colnames(features_all) <- sample_names 

features_all_clean <- features_all %>%
  select(-c(IGF136815, IGF136816, IGF136817))
```

Normalize feature counts and conduct pca
```{r norm_all_consensus_feature_data}
dds <- DESeqDataSetFromMatrix(countData = features_all_clean,
                              colData = meta %>%
                                subset(!grepl("IGF136815|IGF136816|IGF136817", sample)),
                              design = ~ 1) 

# peaks with very low counts 
# dds <- dds[rowSums(counts(dds)) >= 10, ]          
vs  <- vst(dds, blind = TRUE)                    
mat <- assay(vs)

pc <- prcomp(t(mat), center = TRUE, scale. = TRUE)
```

### Percent of variance in feature counts captured by individual PCs

```{r pc_var_plot}
# get variance in each pc
pvar <- (pc$sdev^2)/sum(pc$sdev^2) * 100

pca_var_plot <- ggplot(as.data.frame(pvar) %>%
                         mutate(pc = as.numeric(rownames(.))),
                       aes(y = pvar, x = pc)) +
  geom_col(fill = "#002965") +
  theme_bw() +
  theme(axis.text = element_text(size = 8)) +
  scale_x_continuous(name = "Principle component") +
  scale_y_continuous(name = "Variance (%)")

ggplotly(pca_var_plot)
```

### Percent of variance in features counts attributable to pariticpant variables

```{r variance_plot}
# create annotated df
meta_anno <- AnnotatedDataFrame(data =  meta %>%
                                  subset(!grepl("IGF136815|IGF136816|IGF136817", sample)))

# assign sample names 
sampleNames(meta_anno) <- filtered_stats %>%
  subset(!grepl("IGF136815|IGF136816|IGF136817", sample)) %>%
  pull(sample)

# create expression data set using features and annotated df
eds <- ExpressionSet(assayData = as.matrix(features_all_clean), phenoData = meta_anno)

# create pvca object for plotting - include all vars
pvcaObj <- pvcaBatchAssess(eds, c("sex", "celltype", "age", "pmi", "group"), threshold = 0.1)

var_plot <- ggplot(pvcaObj$dat %>%
                     as.data.frame() %>% 
                     pivot_longer(everything()) %>%
                     mutate(name = pvcaObj$label), 
                   aes(x = reorder(name, -value), 
                       y = value)) +
  geom_col(fill = "#002965") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90,
                                   hjust = 0, 
                                   vjust = 0.5))

ggplotly(var_plot)
```

### Cell type clustering

```{r celltype_clustering}
# combine metadata with pcs
pcs <- data.frame(pc$x[,1:5], 
                  meta %>%
                    subset(!grepl("IGF136815|IGF136816|IGF136817", sample)))

celltype_plot <- ggplot(pcs, 
                        aes(PC1, 
                            PC2, 
                            color = celltype, 
                            shape = group)) +
  geom_point(size=3,
             alpha = 0.8) +
  labs(x=paste0("PC1 (", round(pvar[1],1), "%)"),
       y=paste0("PC2 (", round(pvar[2],1), "%)")) +
  theme_bw() +
  scale_colour_manual(values = c("#00033f", "#e1235c", "#7c91a7"))

ggplotly(celltype_plot)
```

## Consensus features across cell types

```{r }
microglia_tsv_files <- list.files(here("data_out",
                                       "070_peaks",
                                       "078_peaks_feature_counts_celltype"), 
                                  pattern = "*microglia*", 
                                  full.names = TRUE)

oligodendrocyte_tsv_files <- list.files(here("data_out",
                                             "070_peaks",
                                             "078_peaks_feature_counts_celltype"), 
                                        pattern = "*oligodendrocyte*", 
                                        full.names = TRUE)

neuron_tsv_files <- list.files(here("data_out",
                                    "070_peaks",
                                    "078_peaks_feature_counts_celltype"), 
                               pattern = "*neuron*", 
                               full.names = TRUE)

features_microglia <- map_dfc(
  microglia_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = F) %>% 
    select(dplyr::last_col())) 

features_oligodendrocyte <- map_dfc(
  oligodendrocyte_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = F) %>% 
    select(dplyr::last_col())) 

features_neuron <- map_dfc(
  neuron_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = F) %>% 
    select(dplyr::last_col())) 

colnames(features_microglia) <- sub("_microglia_multicov\\.tsv$", 
                                    "", 
                                    basename(microglia_tsv_files))

colnames(features_oligodendrocyte) <- sub("_oligodendrocytes_multicov\\.tsv$", 
                                          "", 
                                          basename(oligodendrocyte_tsv_files))

colnames(features_neuron) <- sub("_neuron_multicov\\.tsv$", 
                                 "", 
                                 basename(neuron_tsv_files))

# remove low frip samples
features_microglia_clean <- features_microglia %>%
  select(!"IGF136817")

features_oligodendrocyte_clean <- features_oligodendrocyte %>%
  select(!"IGF136816")

features_neuron_clean <- features_neuron %>%
  select(!"IGF136815")
```

Normalize microglia feature counts and conduct pca
```{r norm_microglia_consensus_feature_data}
dds_microglia <- DESeqDataSetFromMatrix(countData = features_microglia_clean,
                                        colData = meta %>%
                                          mutate(sample = str_extract(sample, "IGF\\d+")) %>%
                                          subset(sample %in% colnames(features_microglia_clean)),
                                        design = ~ 1) 

# peaks with very low counts 
# dds <- dds[rowSums(counts(dds)) >= 10, ]          
vs_microglia  <- vst(dds_microglia, blind = TRUE)                    
mat_microglia <- assay(vs_microglia)

pc_microglia <- prcomp(t(mat_microglia), center = TRUE, scale. = TRUE)
```


```{r microglia_heat_scree}
pc_microglia_importance <- (pc_microglia$sdev^2) / sum(pc_microglia$sdev^2)

meta_continuous <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  merge(.,
        frip_scores %>%
          select(file, percent),
        by.x = "sample",
        by.y = "file") %>%
  subset(sample %in% colnames(features_microglia_clean)) %>%
  select(age, pmi, percent) %>%
  dplyr::rename(frip = percent)

meta_categorical <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(features_microglia_clean)) %>%
  select(sex, group)

heat_scree_plot(pc_microglia$x, pc_microglia_importance, 8, c(1:8), meta_categorical, meta_continuous)
```


Normalize oligodendrocyte feature counts and conduct pca
```{r norm_oligodendrocyte_consensus_feature_data}
dds_oligodendrocyte <- DESeqDataSetFromMatrix(countData = features_oligodendrocyte_clean,
                                              colData = meta %>%
                                                mutate(sample = str_extract(sample, "IGF\\d+")) %>%
                                                subset(sample %in% colnames(features_oligodendrocyte_clean)),
                                              design = ~ 1) 

# peaks with very low counts 
# dds <- dds[rowSums(counts(dds)) >= 10, ]          
vs_oligodendrocyte  <- vst(dds_oligodendrocyte, blind = TRUE)                    
mat_oligodendrocyte <- assay(vs_oligodendrocyte)

pc_oligodendrocyte <- prcomp(t(mat_oligodendrocyte), center = TRUE, scale. = TRUE)
```


```{r oligodendrocyte_heat_scree}
pc_oligodendrocyte_importance <- (pc_oligodendrocyte$sdev^2) / sum(pc_oligodendrocyte$sdev^2)

meta_continuous <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  merge(.,
        frip_scores %>%
          select(file, percent),
        by.x = "sample",
        by.y = "file") %>%
  subset(sample %in% colnames(features_oligodendrocyte_clean)) %>%
  select(age, pmi, percent) %>%
  dplyr::rename(frip = percent)

meta_categorical <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(features_oligodendrocyte_clean)) %>%
  select(sex, group)

heat_scree_plot(pc_oligodendrocyte$x, pc_oligodendrocyte_importance, 8, c(1:8), meta_categorical, meta_continuous)
```


Normalize neuron feature counts and conduct pca
```{r norm_neuron_consensus_feature_data}
dds_neuron <- DESeqDataSetFromMatrix(countData = features_neuron_clean,
                                     colData = meta %>%
                                       mutate(sample = str_extract(sample, "IGF\\d+")) %>%
                                       subset(sample %in% colnames(features_neuron_clean)),
                                     design = ~ 1) 

# peaks with very low counts 
# dds <- dds[rowSums(counts(dds)) >= 10, ]          
vs_neuron  <- vst(dds_neuron, blind = TRUE)                    
mat_neuron <- assay(vs_neuron)

pc_neuron <- prcomp(t(mat_neuron), center = TRUE, scale. = TRUE)
```

```{r neuron_heat_scree}
pc_neuron_importance <- (pc_neuron$sdev^2) / sum(pc_neuron$sdev^2)

meta_continuous <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  merge(.,
        frip_scores %>%
          select(file, percent),
        by.x = "sample",
        by.y = "file") %>%
  subset(sample %in% colnames(features_neuron_clean)) %>%
  select(age, pmi, percent) %>%
  dplyr::rename(frip = percent)

meta_categorical <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(features_neuron_clean)) %>%
  select(sex, group)

heat_scree_plot(pc_neuron$x, pc_neuron_importance, 8, c(1:8), meta_categorical, meta_continuous)
```

### Percent of variance in cell type feature counts captured by individual PCs

```{r ceclltype_pc_var_plot}
# get variance in each pc
pvar_microglia <- (pc_microglia$sdev^2)/sum(pc_microglia$sdev^2) * 100
pvar_oligodendrocyte <- (pc_oligodendrocyte$sdev^2)/sum(pc_oligodendrocyte$sdev^2) * 100
pvar_neuron <- (pc_neuron$sdev^2)/sum(pc_neuron$sdev^2) * 100

pca_var_celltypes_plot <- data.frame(neuron = data.frame(pvar_neuron),
                                     microglia = data.frame(pvar_microglia),
                                     oligodendrocyte = data.frame(pvar_oligodendrocyte)) %>%
  mutate(pc = as.numeric(rownames(.))) %>%
  pivot_longer(-pc) %>%
  mutate(name = fct_recode(as.factor(name),
                           Microglia = "pvar_microglia",
                           Neurons = "pvar_neuron",
                           Oligodendrocytes = "pvar_oligodendrocyte")) %>%
  ggplot(aes(x = as.character(pc), y = value)) +
  geom_col(fill = "#002965") +
  theme_bw() +
  facet_wrap(~name) +
  scale_y_continuous(name = "Variance (%)") +
  scale_x_discrete(name = "Principal Component (PC)")

ggplotly( pca_var_celltypes_plot)
```





