---
title: "Differential analysis of bulk CUT&Tag data"
author: "Samantha Lee"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load data 

Load required libraries
```{r req_libs}
library(here) # dir calling
library(tidyverse) # data wrangling
library(sva) # surrogate variables
library(DESeq2) # diff analysis
library(ChIPseeker) # peak anno
library(clusterProfiler) # go analysis
library(TxDb.Hsapiens.UCSC.hg38.knownGene) # anno library
library(org.Hs.eg.db) # human annotation
```

Read in participant metadata.
```{r data_in_meta}
meta <- read.csv(here("resources",
                      "metadata",
                      "metadata_participants.csv")) %>%
  mutate(group = ifelse(grepl("PDC", pid), "control", "pd")) %>%
  mutate(sample = str_extract(sample, "IGF\\d+"))
```

Read in FRiP scores. 
```{r data_in_frip_scores}
frip_in <- function(f) {
  read_table(f,
             skip = 0, 
             col_names = T)
}

frip_scores <- map_dfr(list.files(here("data_out",
                                       "070_peaks", 
                                       "072_frip_scores"),
                                  pattern = "^.*\\.txt$",
                                  full.names = TRUE), 
                       frip_in) 
```

Combine metadata with FRiP score.
```{r combine_meta_frip}
meta <- merge(meta,
              frip_scores %>%
                dplyr::select(file, percent) %>%
                dplyr::rename("frip" = percent,
                              "sample" = file),
              by = "sample")
```

Read in celltype-specific feature counts.
```{r data_in_feature_counts}
micro_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*microglia*", 
                              full.names = TRUE)

oligo_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*oligodendrocyte*", 
                              full.names = TRUE)

neuro_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*neuron*", 
                              full.names = TRUE)

micro_feat <- purrr::imap_dfc(
  micro_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)


oligo_feat <- purrr::imap_dfc(
  oligo_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)

neuro_feat <- purrr::imap_dfc(
  neuro_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)

colnames(micro_feat) <- c("chr", "start", "end",
                          sub("_microglia_multicov\\.tsv$", 
                              "", 
                              basename(micro_tsv_files)))

colnames(oligo_feat) <- c("chr", "start", "end",
                          sub("_oligodendrocytes_multicov\\.tsv$", 
                              "", 
                              basename(oligo_tsv_files)))

colnames(neuro_feat) <- c("chr", "start", "end",
                          sub("_neuron_multicov\\.tsv$", 
                              "", 
                              basename(neuro_tsv_files)))

# remove low frip samples
micro_feat_clean <- micro_feat %>%
  dplyr::select(!"IGF136817")

oligo_feat_clean <- oligo_feat %>%
  dplyr::select(!"IGF136816")

neuro_feat_clean <- neuro_feat %>%
  dplyr::select(!"IGF136815")

# remove uneeded files 
rm(micro_tsv_files,
   oligo_tsv_files,
   neuro_tsv_files,
   micro_feat,
   oligo_feat,
   neuro_feat,
   frip_scores)
```

Subset metadata for each cell type
```{r meta_celltype}
micro_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(micro_feat_clean))

oligo_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(oligo_feat_clean))

neuro_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(neuro_feat_clean))
```

## Surrogate variable analysis (SVA)

Attempt to remove biological and technical variation using surrogate variables (SVs). Use instead of raw metadata to reduce covariates in linear regressions for DESeq2. 

### SVA microglia 

Calculate surrogate variables (SVs) for microglia.
```{r sva_microglia}
# create model matrix 
micro_mod = model.matrix(~as.factor(group) + pmi,
                         data = micro_meta)
# create null matrix
micro_mod0 = model.matrix(~1,
                          data=micro_meta)

# estimate num svs req
micro_nsv = num.sv(micro_feat_clean[,-c(1:3)],
                   micro_mod)

# 1 sv per above but calculate five to see correlations
micro_svobj = sva(as.matrix(micro_feat_clean[,-c(1:3)]),
                  micro_mod,
                  micro_mod0,
                  n.sv=5)
```

Examine correlation between microglia SVs and metadata
```{r micro_sv_meta_corr}
# bind meta + svs
micro_meta_svs <- cbind(micro_meta, micro_svobj$sv)
# rename sv cols
colnames(micro_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
micro_meta_corr <- micro_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

micro_corr <- round(cor(micro_meta_corr),2)
micro_corr_long <- reshape2::melt(micro_corr)

ggplot(data = micro_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```


### SVA oligodendrocytes 

Calculate surrogate variables (SVs) for oligoglia.
```{r sva_oligoglia}
# create model matrix 
oligo_mod = model.matrix(~as.factor(group),
                         data = oligo_meta)
# create null matrix
oligo_mod0 = model.matrix(~1,
                          data=oligo_meta)

# estimate num svs req
oligo_nsv = num.sv(oligo_feat_clean[,-c(1:3)],
                   oligo_mod)

# 1 sv per above but calculate five to see correlations
oligo_svobj = sva(as.matrix(oligo_feat_clean[,-c(1:3)]),
                  oligo_mod,
                  oligo_mod0,
                  n.sv=5)
```

Examine correlation between oligoglia SVs and metadata
```{r oligo_sv_meta_corr}
# bind meta + svs
oligo_meta_svs <- cbind(oligo_meta, oligo_svobj$sv)
# rename sv cols
colnames(oligo_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
oligo_meta_corr <- oligo_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

oligo_corr <- round(cor(oligo_meta_corr),2)
oligo_corr_long <- reshape2::melt(oligo_corr)

ggplot(data = oligo_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```


### SVA neurons 

Calculate surrogate variables (SVs) for neuroglia.
```{r sva_neuroglia}
# create model matrix 
neuro_mod = model.matrix(~as.factor(group),
                         data = neuro_meta)
# create null matrix
neuro_mod0 = model.matrix(~1,
                          data=neuro_meta)

# estimate num svs req
neuro_nsv = num.sv(neuro_feat_clean[,-c(1:3)],
                   neuro_mod)

# 1 sv per above but calculate five to see correlations
neuro_svobj = sva(as.matrix(neuro_feat_clean[,-c(1:3)]),
                  neuro_mod,
                  neuro_mod0,
                  n.sv=5)
```

Examine correlation between neuroglia SVs and metadata
```{r neuro_sv_meta_corr}
# bind meta + svs together
neuro_meta_svs <- cbind(neuro_meta, neuro_svobj$sv)
# rename sv cols
colnames(neuro_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
neuro_meta_corr <- neuro_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

neuro_corr <- round(cor(neuro_meta_corr),2)
neuro_corr_long <- reshape2::melt(neuro_corr)

ggplot(data = neuro_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```

## Differential analysis

### Differential analysis of microglia

```{r micro_diff}
# only keep rows with 5 or more counts
micro_feat_clean_hc <- micro_feat_clean[which(rowSums(micro_feat_clean[,-c(1:3)]) >= 5), ]

# make deseq2 object
micro_dds = DESeqDataSetFromMatrix(countData = micro_feat_clean_hc[,-c(1:3)],
                                   colData = micro_meta_svs %>%
                                     mutate(group = as.factor(group)),
                                   design = ~group + sv1)

# diff analysis
micro_dds_diff = DESeq(micro_dds, 
                       test = "Wald", 
                       fitType = "local", 
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
micro_dds_norm_counts = counts(micro_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
micro_dds_res = results(micro_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
resultsNames(micro_dds_diff)

# fold change shrinkage
micro_dds_res_shr <- lfcShrink(micro_dds_diff, 
                               coef = "group_pd_vs_control", 
                               type = "apeglm")

# make results df
micro_dds_res_df <- data.frame(chr = micro_feat_clean$chr,
                               start = micro_feat_clean$start,
                               end = micro_feat_clean$end,
                               baseMean = micro_dds_res$baseMean,
                               lfc = micro_dds_res$log2FoldChange,
                               lfcSE = micro_dds_res$lfcSE,
                               lfc_shr = micro_dds_res_shr$log2FoldChange,
                               stat = micro_dds_res$stat,
                               pvalue = micro_dds_res$pvalue,
                               padj = micro_dds_res$padj) 

# get only sig hyper acetylated peaks
micro_dds_res_df_hyper <- micro_dds_res_df %>%
  subset(padj < 0.05 & lfc > 0)

# get only sig hypoacetylated peaks
micro_dds_res_df_hypo <- micro_dds_res_df %>%
  subset(padj < 0.05 & lfc < 0)
```

```{r micro_qqplot}
qq_p_gg <- function(p, ci = 0.95) {
  
  p <- p[is.finite(p) & !is.na(p)]
  p[p <= 0] <- .Machine$double.xmin
  p[p > 1] <- 1
  
  o <- sort(p)
  n <- length(o); i <- seq_len(n)
  e <- (i - 0.5) / n
  alpha <- (1 - ci) / 2
  lo <- qbeta(alpha, i, n - i + 1)
  hi <- qbeta(1 - alpha, i, n - i + 1)
  
  df <- data.frame(
    exp = -log10(e),
    obs = -log10(o),
    lower = -log10(lo),
    upper = -log10(hi)
  )
  
  ggplot(df, aes(exp, obs)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) +
    geom_abline(intercept = 0, slope = 1, linetype = 2) +
    geom_point(size = 0.8, alpha = 0.6, color = "#00033f") +
    coord_equal() +
    labs(x = "Expected -log10(p)", y = "Observed -log10(p)") +
    theme_minimal()
}

qq_p_gg(micro_dds_res_df$pvalue)
```

```{r micro_volcano}
ggplot(micro_dds_res_df %>%
         mutate(color = ifelse(padj < 0.05 & abs(lfc) >=2, 
                               "#e1235c", 
                               "#00033f")), 
       aes(x = lfc, 
           y = -log(padj),
           color = color)) +
  geom_point() +
  theme_bw() +
  scale_color_manual(values = c( "#00033f", "#e1235c"))
```

```{r micro_go_terms_hyper}
# convert peaks to granges obj
micro_dds_res_df_hyper_gr <- GRanges(seqnames = micro_dds_res_df_hyper$chr,
                                     ranges   = IRanges(as.numeric(micro_dds_res_df_hyper$start), 
                                                        as.numeric(micro_dds_res_df_hyper$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(micro_dds_res_df_hyper_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
micro_dds_res_df_hyper_anno <- annotatePeak(micro_dds_res_df_hyper_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
micro_dds_res_df_hyper <- merge(micro_dds_res_df_hyper,
                               as.data.frame(micro_dds_res_df_hyper_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

micro_hyper_genes <- unique(na.omit(as.data.frame(micro_dds_res_df_hyper_anno)$ENSEMBL))

micro_hyper_go <- enrichGO(gene=micro_hyper_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(micro_hyper_go)
```

```{r micro_go_terms_hypo}
# convert peaks to granges obj
micro_dds_res_df_hypo_gr <- GRanges(seqnames = micro_dds_res_df_hypo$chr,
                                     ranges   = IRanges(as.numeric(micro_dds_res_df_hypo$start), 
                                                        as.numeric(micro_dds_res_df_hypo$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(micro_dds_res_df_hypo_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
micro_dds_res_df_hypo_anno <- annotatePeak(micro_dds_res_df_hypo_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
micro_dds_res_df_hypo <- merge(micro_dds_res_df_hypo,
                               as.data.frame(micro_dds_res_df_hypo_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

micro_hypo_genes <- unique(na.omit(as.data.frame(micro_dds_res_df_hypo_anno)$ENSEMBL))

micro_hypo_go <- enrichGO(gene=micro_hypo_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(micro_hypo_go)
```

### Oligodendrocyte diff 

```{r oligo_diff}
# only keep rows with 5 or more counts
oligo_feat_clean_hc <- oligo_feat_clean[which(rowSums(oligo_feat_clean[,-c(1:3)]) >= 5), ]

# make deseq2 object
oligo_dds = DESeqDataSetFromMatrix(countData = oligo_feat_clean_hc[,-c(1:3)],
                                   colData = oligo_meta_svs %>%
                                     mutate(group = as.factor(group)),
                                   design = ~group + sv1)

# diff analysis
oligo_dds_diff = DESeq(oligo_dds, 
                       test = "Wald", 
                       fitType = "local", 
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
oligo_dds_norm_counts = counts(oligo_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
oligo_dds_res = results(oligo_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
resultsNames(oligo_dds_diff)

# fold change shrinkage
oligo_dds_res_shr <- lfcShrink(oligo_dds_diff, 
                               coef = "group_pd_vs_control", 
                               type = "apeglm")

# make results df
oligo_dds_res_df <- data.frame(chr = oligo_feat_clean$chr,
                               start = oligo_feat_clean$start,
                               end = oligo_feat_clean$end,
                               baseMean = oligo_dds_res$baseMean,
                               lfc = oligo_dds_res$log2FoldChange,
                               lfcSE = oligo_dds_res$lfcSE,
                               lfc_shr = oligo_dds_res_shr$log2FoldChange,
                               stat = oligo_dds_res$stat,
                               pvalue = oligo_dds_res$pvalue,
                               padj = oligo_dds_res$padj) 

# get only sig hyper acetylated peaks
oligo_dds_res_df_hyper <- oligo_dds_res_df %>%
  subset(padj < 0.05 & lfc > 0)

# get only sig hypoacetylated peaks
oligo_dds_res_df_hypo <- oligo_dds_res_df %>%
  subset(padj < 0.05 & lfc < 0)
```

```{r oligo_qqplot}
qq_p_gg(oligo_dds_res_df$pvalue)
```

```{r oligo_volcano}
ggplot(oligo_dds_res_df %>%
         mutate(color = ifelse(padj < 0.05 & abs(lfc) >=2, 
                               "#e1235c", 
                               "#00033f")), 
       aes(x = lfc, 
           y = -log(padj),
           color = color)) +
  geom_point() +
  theme_bw() +
  scale_color_manual(values = c( "#00033f", "#e1235c"))
```

```{r oligo_go_terms_hyper}
# convert peaks to granges obj
oligo_dds_res_df_hyper_gr <- GRanges(seqnames = oligo_dds_res_df_hyper$chr,
                                     ranges   = IRanges(as.numeric(oligo_dds_res_df_hyper$start), 
                                                        as.numeric(oligo_dds_res_df_hyper$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(oligo_dds_res_df_hyper_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
oligo_dds_res_df_hyper_anno <- annotatePeak(oligo_dds_res_df_hyper_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
oligo_dds_res_df_hyper <- merge(oligo_dds_res_df_hyper,
                               as.data.frame(oligo_dds_res_df_hyper_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

oligo_hyper_genes <- unique(na.omit(as.data.frame(oligo_dds_res_df_hyper_anno)$ENSEMBL))

oligo_hyper_go <- enrichGO(gene=oligo_hyper_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(oligo_hyper_go)
```

```{r oligo_go_terms_hypo}
# convert peaks to granges obj
oligo_dds_res_df_hypo_gr <- GRanges(seqnames = oligo_dds_res_df_hypo$chr,
                                     ranges   = IRanges(as.numeric(oligo_dds_res_df_hypo$start), 
                                                        as.numeric(oligo_dds_res_df_hypo$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(oligo_dds_res_df_hypo_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
oligo_dds_res_df_hypo_anno <- annotatePeak(oligo_dds_res_df_hypo_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
oligo_dds_res_df_hypo <- merge(oligo_dds_res_df_hypo,
                               as.data.frame(oligo_dds_res_df_hypo_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

oligo_hypo_genes <- unique(na.omit(as.data.frame(oligo_dds_res_df_hypo_anno)$ENSEMBL))

oligo_hypo_go <- enrichGO(gene=oligo_hypo_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(oligo_hypo_go)
```

### Differential analysis of neurons

```{r neuro_diff}
# only keep rows with 5 or more counts
neuro_feat_clean_hc <- neuro_feat_clean[which(rowSums(neuro_feat_clean[,-c(1:3)]) >= 5), ]

# make deseq2 object
neuro_dds = DESeqDataSetFromMatrix(countData = neuro_feat_clean_hc[,-c(1:3)],
                                   colData = neuro_meta_svs %>%
                                     mutate(group = as.factor(group)),
                                   design = ~group + sv1)

# diff analysis
neuro_dds_diff = DESeq(neuro_dds, 
                       test = "Wald", 
                       fitType = "local", 
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
neuro_dds_norm_counts = counts(neuro_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
neuro_dds_res = results(neuro_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
resultsNames(neuro_dds_diff)

# fold change shrinkage
neuro_dds_res_shr <- lfcShrink(neuro_dds_diff, 
                               coef = "group_pd_vs_control", 
                               type = "apeglm")

# make results df
neuro_dds_res_df <- data.frame(chr = neuro_feat_clean$chr,
                               start = neuro_feat_clean$start,
                               end = neuro_feat_clean$end,
                               baseMean = neuro_dds_res$baseMean,
                               lfc = neuro_dds_res$log2FoldChange,
                               lfcSE = neuro_dds_res$lfcSE,
                               lfc_shr = neuro_dds_res_shr$log2FoldChange,
                               stat = neuro_dds_res$stat,
                               pvalue = neuro_dds_res$pvalue,
                               padj = neuro_dds_res$padj) 

# get only sig hyper acetylated peaks
neuro_dds_res_df_hyper <- neuro_dds_res_df %>%
  subset(padj < 0.30 & lfc > 0)

# get only sig hypoacetylated peaks
neuro_dds_res_df_hypo <- neuro_dds_res_df %>%
  subset(padj < 0.30 & lfc < 0)
```

```{r neuro_qqplot}
qq_p_gg(neuro_dds_res_df$pvalue)
```

```{r neuro_volcano}
ggplot(neuro_dds_res_df %>%
         mutate(color = ifelse(padj < 0.05 & abs(lfc) >=2, 
                               "#e1235c", 
                               "#00033f")), 
       aes(x = lfc, 
           y = -log(padj),
           color = color)) +
  geom_point() +
  theme_bw() +
  scale_color_manual(values = c( "#00033f", "#e1235c"))
```

```{r neuro_go_terms_hyper}
# convert peaks to granges obj
neuro_dds_res_df_hyper_gr <- GRanges(seqnames = neuro_dds_res_df_hyper$chr,
                                     ranges   = IRanges(as.numeric(neuro_dds_res_df_hyper$start), 
                                                        as.numeric(neuro_dds_res_df_hyper$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(neuro_dds_res_df_hyper_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
neuro_dds_res_df_hyper_anno <- annotatePeak(neuro_dds_res_df_hyper_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
neuro_dds_res_df_hyper <- merge(neuro_dds_res_df_hyper,
                               as.data.frame(neuro_dds_res_df_hyper_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

neuro_hyper_genes <- unique(na.omit(as.data.frame(neuro_dds_res_df_hyper_anno)$ENSEMBL))

neuro_hyper_go <- enrichGO(gene=neuro_hyper_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(neuro_hyper_go)
```

```{r neuro_go_terms_hypo}
# convert peaks to granges obj
neuro_dds_res_df_hypo_gr <- GRanges(seqnames = neuro_dds_res_df_hypo$chr,
                                     ranges   = IRanges(as.numeric(neuro_dds_res_df_hypo$start), 
                                                        as.numeric(neuro_dds_res_df_hypo$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(neuro_dds_res_df_hypo_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
neuro_dds_res_df_hypo_anno <- annotatePeak(neuro_dds_res_df_hypo_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
neuro_dds_res_df_hypo <- merge(neuro_dds_res_df_hypo,
                               as.data.frame(neuro_dds_res_df_hypo_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

neuro_hypo_genes <- unique(na.omit(as.data.frame(neuro_dds_res_df_hypo_anno)$ENSEMBL))

neuro_hypo_go <- enrichGO(gene=neuro_hypo_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(neuro_hypo_go)
```

### Output size factors

```{r size factors}
# get size factors
micro_size_factors <- sizeFactors(micro_dds_diff) %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    merge(.,
              micro_meta_svs %>% 
              dplyr::select(sample, celltype, group),
              by.x = "rowname", by.y= "sample")

oligo_size_factors <- sizeFactors(oligo_dds_diff) %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    merge(.,
              oligo_meta_svs %>% 
              dplyr::select(sample, celltype, group),
              by.x = "rowname", by.y= "sample")

neuro_size_factors <- sizeFactors(neuro_dds_diff) %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    merge(.,
              neuro_meta_svs %>% 
              dplyr::select(sample, celltype, group),
              by.x = "rowname", by.y= "sample")

size_factors <- rbind(micro_size_factors,
                      oligo_size_factors,
                      neuro_size_factors) %>%
  as.data.frame() 

write_tsv(size_factors,
          here("data_out",
               "080_differential_analysis",
               "size_factors.tsv"),
          col_names = FALSE)
```


