---
title: "Differential analysis of bulk CUT&Tag data"
author: "Samantha Lee"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load data 

Load required libraries
```{r req_libs}
library(here) # dir calling
library(tidyverse) # data wrangling
library(sva) # surrogate variables
library(DESeq2) # diff analysis
library(edgeR) # diff analysis 
library(ChIPseeker) # peak anno
library(clusterProfiler) # go analysis
library(TxDb.Hsapiens.UCSC.hg38.knownGene) # anno library
library(org.Hs.eg.db) # human annotation
```

Read in participant metadata.
```{r data_in_meta}
meta <- read.csv(here("resources",
                      "metadata",
                      "metadata_participants.csv")) %>%
  mutate(group = ifelse(grepl("PDC", pid), "control", "pd")) %>%
  mutate(sample = str_extract(sample, "IGF\\d+"))
```

Read in FRiP scores. 
```{r data_in_frip_scores}
frip_in <- function(f) {
  read_table(f,
             skip = 0, 
             col_names = T)
}

frip_scores <- map_dfr(list.files(here("data_out",
                                       "070_peaks", 
                                       "072_frip_scores"),
                                  pattern = "^.*\\.txt$",
                                  full.names = TRUE), 
                       frip_in) 
```

Combine metadata with FRiP score.
```{r combine_meta_frip}
meta <- merge(meta,
              frip_scores %>%
                dplyr::select(file, percent) %>%
                dplyr::rename("frip" = percent,
                              "sample" = file),
              by = "sample")
```

Read in celltype-specific feature counts.
```{r data_in_feature_counts}
micro_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*microglia*", 
                              full.names = TRUE)

oligo_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*oligodendrocyte*", 
                              full.names = TRUE)

neuro_tsv_files <- list.files(here("data_out",
                                   "070_peaks",
                                   "078_peaks_feature_counts_celltype"), 
                              pattern = "*neuron*", 
                              full.names = TRUE)

micro_feat <- purrr::imap_dfc(
  micro_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)


oligo_feat <- purrr::imap_dfc(
  oligo_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)

neuro_feat <- purrr::imap_dfc(
  neuro_tsv_files,
  ~ read_tsv(.x, show_col_types = FALSE, col_names = FALSE) %>% 
    { if (.y == 1) dplyr::select(., 1:3, dplyr::last_col())
      else          dplyr::select(., dplyr::last_col()) }
)

colnames(micro_feat) <- c("chr", "start", "end",
                          sub("_microglia_multicov\\.tsv$", 
                              "", 
                              basename(micro_tsv_files)))

colnames(oligo_feat) <- c("chr", "start", "end",
                          sub("_oligodendrocytes_multicov\\.tsv$", 
                              "", 
                              basename(oligo_tsv_files)))

colnames(neuro_feat) <- c("chr", "start", "end",
                          sub("_neuron_multicov\\.tsv$", 
                              "", 
                              basename(neuro_tsv_files)))

# remove low frip samples
micro_feat_clean <- micro_feat %>%
  dplyr::select(!"IGF136817")

oligo_feat_clean <- oligo_feat %>%
  dplyr::select(!"IGF136816")

neuro_feat_clean <- neuro_feat %>%
  dplyr::select(!"IGF136815")

# remove uneeded files 
rm(micro_tsv_files,
   oligo_tsv_files,
   neuro_tsv_files,
   micro_feat,
   oligo_feat,
   neuro_feat,
   frip_scores)
```

Subset metadata for each cell type
```{r meta_celltype}
micro_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(micro_feat_clean))

oligo_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(oligo_feat_clean))

neuro_meta <- meta %>%
  mutate(sample = str_extract(sample, "IGF\\d+")) %>%
  subset(sample %in% colnames(neuro_feat_clean))
```

## Annotate peaks for each cell type

```{r micro_peak_anno}
# convert peaks to granges obj
micro_feat_gr  <- GRanges(seqnames = micro_feat_clean$chr,
                          ranges   = IRanges(as.numeric(micro_feat_clean$start), 
                                             as.numeric(micro_feat_clean$end)),
                          strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(micro_feat_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
micro_feat_anno <- annotatePeak(micro_feat_gr, 
                                TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                tssRegion=c(-3000,3000), 
                                annoDb="org.Hs.eg.db")

# merge annotate to res df
micro_feat_clean_anno <- merge(micro_feat_clean,
                               as.data.frame(micro_feat_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))
```

```{r oligo_peak_anno}
# convert peaks to granges obj
oligo_feat_gr  <- GRanges(seqnames = oligo_feat_clean$chr,
                          ranges   = IRanges(as.numeric(oligo_feat_clean$start), 
                                             as.numeric(oligo_feat_clean$end)),
                          strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(oligo_feat_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
oligo_feat_anno <- annotatePeak(oligo_feat_gr, 
                                TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                tssRegion=c(-3000,3000), 
                                annoDb="org.Hs.eg.db")

# merge annotate to res df
oligo_feat_clean_anno <- merge(oligo_feat_clean,
                               as.data.frame(oligo_feat_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))
```

```{r neuro_peak_anno}
# convert peaks to granges obj
neuro_feat_gr  <- GRanges(seqnames = neuro_feat_clean$chr,
                          ranges   = IRanges(as.numeric(neuro_feat_clean$start), 
                                             as.numeric(neuro_feat_clean$end)),
                          strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(neuro_feat_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
neuro_feat_anno <- annotatePeak(neuro_feat_gr, 
                                TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                tssRegion=c(-3000,3000), 
                                annoDb="org.Hs.eg.db")

# merge annotate to res df
neuro_feat_clean_anno <- merge(neuro_feat_clean,
                               as.data.frame(neuro_feat_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))
```

## Covariate distributions

```{r covariates}
sex_plot <- ggplot(neuro_meta, 
                   aes(x=group, fill = sex)) +
  geom_bar() +
  theme_bw()
sex_plot
# sex included regardless 

age_plot <- ggplot(neuro_meta, 
                   aes(x=group,
                       y = age, 
                       fill = group)) +
  geom_boxplot() +
  geom_point() +
  theme_bw()
age_plot

age_ttest <- t.test(age ~ group, data=neuro_meta)
age_ttest # p = 0.03

agesex_plot <- ggplot(neuro_meta, 
                   aes(x=sex,
                       y = age, 
                       fill = sex)) +
  geom_boxplot() +
  geom_point() +
  theme_bw()
agesex_plot

agesex_ttest <- t.test(age ~ sex, data=neuro_meta)
agesex_ttest # p = 0.07064

frip_plot <- ggplot(neuro_meta, 
                    aes(x=group,
                        y = frip, 
                        fill = group)) +
  geom_boxplot() +
  geom_point() +
  theme_bw()
frip_plot

frip_ttest <- t.test(frip ~ group, data=neuro_meta)
frip_ttest # p = 0.27

pmi_plot <- ggplot(neuro_meta, 
                   aes(x=group,
                       y = pmi, 
                       fill = group)) +
  geom_boxplot() +
  geom_point() +
  theme_bw()
pmi_plot

pmi_ttest <- t.test(pmi ~ group, data=neuro_meta)
pmi_ttest # p = 0.66
```

## Surrogate variable analysis (SVA)

Attempt to remove biological and technical variation using surrogate variables (SVs). Use instead of raw metadata to reduce covariates in linear regressions for DESeq2. 

### SVA microglia 

Calculate surrogate variables (SVs) for microglia.
```{r sva_microglia}
# create model matrix 
mod = model.matrix(~as.factor(group) + frip + sex,
                         data = micro_meta)
# create null matrix
mod0 = model.matrix(~1,
                          data=micro_meta)

# estimate num svs req
#micro_nsv = num.sv(micro_feat_clean[,-c(1:3)],
#                   micro_mod)

# 1 sv per above but calculate five to see correlations
micro_svobj = sva(as.matrix(micro_feat_clean[,-c(1:3)]),
                  micro_mod,
                  micro_mod0,
                  n.sv=5)
```

Examine correlation between microglia SVs and metadata
```{r micro_sv_meta_corr}
# bind meta + svs
micro_meta_svs <- cbind(micro_meta, micro_svobj$sv)
# rename sv cols
colnames(micro_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
micro_meta_corr <- micro_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

micro_corr <- round(cor(micro_meta_corr),2)
micro_corr_long <- reshape2::melt(micro_corr)

ggplot(data = micro_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```

```{r micro_sva_scree}
data.frame(SV = factor(colnames(micro_svobj$sv), levels = colnames(micro_svobj$sv)),
                 PropVar = prop_var, CumVar = cum_prop)
ggplot(df, aes(SV, PropVar*100)) +
  geom_col() +
  geom_point(aes(y = CumVar*100), size = 2) +
  geom_line(aes(y = CumVar*100, group = 1)) +
  labs(y = "% variance (PCA on SVs)", x = "Surrogate variable",
       title = "Scree plot of SVs (PCA on SV matrix)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### SVA oligodendrocytes 

Calculate surrogate variables (SVs) for oligoglia.
```{r sva_oligoglia}
# create model matrix 
oligo_mod = model.matrix(~as.factor(group),
                         data = oligo_meta)
# create null matrix
oligo_mod0 = model.matrix(~1,
                          data=oligo_meta)

# estimate num svs req
#oligo_nsv = num.sv(oligo_feat_clean[,-c(1:3)],
#                   oligo_mod)

# 1 sv per above but calculate five to see correlations
oligo_svobj = sva(as.matrix(oligo_feat_clean[,-c(1:3)]),
                  oligo_mod,
                  oligo_mod0,
                  n.sv=5)
```

Examine correlation between oligoglia SVs and metadata
```{r oligo_sv_meta_corr}
# bind meta + svs
oligo_meta_svs <- cbind(oligo_meta, oligo_svobj$sv)
# rename sv cols
colnames(oligo_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
oligo_meta_corr <- oligo_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

oligo_corr <- round(cor(oligo_meta_corr),2)
oligo_corr_long <- reshape2::melt(oligo_corr)

ggplot(data = oligo_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```


### SVA neurons 

Calculate surrogate variables (SVs) for neuroglia.
```{r sva_neuroglia}
# create model matrix 
neuro_mod = model.matrix(~as.factor(group),
                         data = neuro_meta)
# create null matrix
neuro_mod0 = model.matrix(~1,
                          data=neuro_meta)

# estimate num svs req
#neuro_nsv = num.sv(neuro_feat_clean[,-c(1:3)],
#                   neuro_mod)

# 1 sv per above but calculate five to see correlations
neuro_svobj = sva(as.matrix(neuro_feat_clean[,-c(1:3)]),
                  neuro_mod,
                  neuro_mod0,
                  n.sv=5)
```

Examine correlation between neuroglia SVs and metadata
```{r neuro_sv_meta_corr}
# bind meta + svs together
neuro_meta_svs <- cbind(neuro_meta, neuro_svobj$sv)
# rename sv cols
colnames(neuro_meta_svs)[9:13] <- c("sv1", "sv2", "sv3", "sv4", "sv5")

# convert all var to num for corr
neuro_meta_corr <- neuro_meta_svs %>%
  dplyr::select(-c(sample, pid, celltype)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(where(is.factor), as.numeric))

neuro_corr <- round(cor(neuro_meta_corr),2)
neuro_corr_long <- reshape2::melt(neuro_corr)

ggplot(data = neuro_corr_long, 
       aes(x=Var1, 
           y=Var2, 
           fill=value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue",
                       high = "red",
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1,1), 
                       space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  geom_text(aes(Var2, 
                Var1,
                label = as.character(value)), 
            color = "black",
            size = 4) 
```

SVs do not adequately capture variation related to age - correct directly for age and sex. 

## Differential analysis

### Differential analysis of microglia

```{r micro_diff_filter}
# remove sex chromosomes
micro_feat_clean_noxy <- micro_feat_clean_anno %>%
  subset(!chr %in% c("X", "Y"))

# separate coordinates and counts
micro_peak_coords <- micro_feat_clean_noxy[, c(1:3,13:27)]
micro_counts <- as.matrix(micro_feat_clean_noxy[, -c(1:3,13:27)])

# recode group as factor
# set control as base level
micro_meta_svs <- micro_meta_svs %>%
  mutate(group = relevel(factor(group), 
                         ref = "control")) 

# create edgeR dgelist for filtering
micro_dgelist <- DGEList(counts = micro_counts, 
                         genes = micro_peak_coords,
                         samples = micro_meta_svs)

# replicate-aware filtering - min count 5 
micro_keep <- filterByExpr(micro_dgelist, 
                           group = micro_dgelist$samples$group, 
                           min.count = 5,
                           min.total.count= 50) 

sum(micro_keep) # 70270

# remove those with <5 from count and peaks
micro_peak_coords_clean <- micro_peak_coords[micro_keep, ]
micro_counts_clean <- micro_counts[micro_keep, ]
```

```{r micro_diff_deseq2}
# make deseq2 object
micro_dds = DESeqDataSetFromMatrix(countData = micro_counts_clean,
                                   colData = micro_meta_svs,
                                   design = ~ group + sex + sv1)

# diff analysis using deseq2
micro_dds_diff = DESeq(micro_dds, 
                       test = "Wald", 
                       fitType = "local",
                      # reduced = ~ sv1, # lrt only
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
micro_dds_norm_counts = counts(micro_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
micro_dds_res = results(micro_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
# resultsNames(micro_dds_diff)

# # fold change shrinkage
# micro_dds_res_shr <- lfcShrink(micro_dds_diff, 
#                                coef = "group_pd_vs_control", 
#                                type = "apeglm")

# make results df
micro_dds_res_df <- data.frame(chr = micro_peak_coords_clean$chr,
                               start = micro_peak_coords_clean$start,
                               end = micro_peak_coords_clean$end,
                               baseMean = micro_dds_res$baseMean,
                               lfc = micro_dds_res$log2FoldChange,
                               lfcSE = micro_dds_res$lfcSE,
                              # lfc_shr = micro_dds_res_shr$log2FoldChange,
                               stat = micro_dds_res$stat,
                               pvalue = micro_dds_res$pvalue,
                               padj = micro_dds_res$padj) 

# get only sig hyper acetylated peaks
micro_dds_res_df_hyper <- micro_dds_res_df %>%
  subset(padj < 0.05 & lfc > 1.5)

dim(micro_dds_res_df_hyper)

# get only sig hypoacetylated peaks
micro_dds_res_df_hypo <- micro_dds_res_df %>%
  subset(padj < 0.05 & lfc < -1.5)

nrow(micro_dds_res_df_hypo)
```

```{r micro_qqplot}
qq_p_gg <- function(p, ci = 0.95) {
  
  p <- p[is.finite(p) & !is.na(p)]
  p[p <= 0] <- .Machine$double.xmin
  p[p > 1] <- 1
  
  o <- sort(p)
  n <- length(o); i <- seq_len(n)
  e <- (i - 0.5) / n
  alpha <- (1 - ci) / 2
  lo <- qbeta(alpha, i, n - i + 1)
  hi <- qbeta(1 - alpha, i, n - i + 1)
  
  df <- data.frame(
    exp = -log10(e),
    obs = -log10(o),
    lower = -log10(lo),
    upper = -log10(hi)
  )
  
  ggplot(df, aes(exp, obs)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15) +
    geom_abline(intercept = 0, slope = 1, linetype = 2) +
    geom_point(size = 0.8, alpha = 0.6, color = "#00033f") +
    coord_equal() +
    labs(x = "Expected -log10(p)", y = "Observed -log10(p)") +
    theme_minimal()
}

lambda_from_p <- function(p, df = 1, na.rm = TRUE) {
  p <- p[is.finite(p)]
  if (na.rm) p <- p[!is.na(p)]
  chisq_obs <- qchisq(1 - p, df = df)
  median(chisq_obs) / qchisq(0.5, df = df)
}

qq_p_gg(micro_dds_res_df$pvalue)
lambda_from_p(micro_dds_res_df$pvalue)
```

```{r micro_volcano}
ggplot(micro_dds_res_df, 
       aes(x = lfc, 
           y = -log(pvalue))) +
  geom_point() +
  theme_bw() 
```

```{r micro_go_terms_hyper}
# convert peaks to granges obj
micro_dds_res_df_hyper_gr <- GRanges(seqnames = micro_dds_res_df_hyper$chr,
                                     ranges   = IRanges(as.numeric(micro_dds_res_df_hyper$start), 
                                                        as.numeric(micro_dds_res_df_hyper$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(micro_dds_res_df_hyper_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
micro_dds_res_df_hyper_anno <- annotatePeak(micro_dds_res_df_hyper_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
micro_dds_res_df_hyper <- merge(micro_dds_res_df_hyper,
                                as.data.frame(micro_dds_res_df_hyper_anno@anno) %>%
                                  mutate(chr = str_sub(seqnames, 4, 6)),
                                by = c("chr", "start", "end"))

micro_hyper_genes <- unique(na.omit(as.data.frame(micro_dds_res_df_hyper_anno)$ENSEMBL))

micro_hyper_go <- enrichGO(gene=micro_hyper_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(micro_hyper_go)
```

```{r micro_go_terms_hypo}
# convert peaks to granges obj
micro_dds_res_df_hypo_gr <- GRanges(seqnames = micro_dds_res_df_hypo$chr,
                                    ranges   = IRanges(as.numeric(micro_dds_res_df_hypo$start), 
                                                       as.numeric(micro_dds_res_df_hypo$end)),
                                    strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(micro_dds_res_df_hypo_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
micro_dds_res_df_hypo_anno <- annotatePeak(micro_dds_res_df_hypo_gr, 
                                           TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                           tssRegion=c(-3000,3000), 
                                           annoDb="org.Hs.eg.db")

# merge annotate to res df
micro_dds_res_df_hypo <- merge(micro_dds_res_df_hypo,
                               as.data.frame(micro_dds_res_df_hypo_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

micro_hypo_genes <- unique(na.omit(as.data.frame(micro_dds_res_df_hypo_anno)$ENSEMBL))

micro_hypo_go <- enrichGO(gene=micro_hypo_genes, 
                          OrgDb=org.Hs.eg.db, 
                          keyType="ENSEMBL",
                          ont="BP", 
                          #pvalueCutoff=0.01, 
                          qvalueCutoff=0.05)
dotplot(micro_hypo_go)
```






### Differential analysis of oligodendrocytes

```{r oligo_diff_filter}
# remove sex chromosomes
oligo_feat_clean_noxy <- oligo_feat_clean_anno %>%
  subset(!chr %in% c("X", "Y"))

# separate coordinates and counts
oligo_peak_coords <- oligo_feat_clean_noxy[, c(1:3, 13:27)]
oligo_counts <- as.matrix(oligo_feat_clean_noxy[, -c(1:3, 13:27)])

# recode group as factor
# set control as base level
oligo_meta_svs <- oligo_meta_svs %>%
  mutate(group = relevel(factor(group), 
                         ref = "control")) 

# create edgeR dgelist for filtering
oligo_dgelist <- DGEList(counts = oligo_counts, 
                         samples = oligo_meta_svs)

# replicate-aware filtering - min count 5 
oligo_keep <- filterByExpr(oligo_dgelist, 
                           group = oligo_dgelist$samples$group, 
                           min.count = 5, 
                           min.total.count=50) # smallest group (control) has 4 samples

sum(oligo_keep) # 76711

# remove those with <5 from count and peaks
oligo_peak_coords_clean <- oligo_peak_coords[oligo_keep, ]
oligo_counts_clean <- oligo_counts[oligo_keep, ]
```

```{r oligo_diff_deseq2}
# make deseq2 object
oligo_dds = DESeqDataSetFromMatrix(countData = oligo_counts_clean,
                                   colData = oligo_meta_svs,
                                   design = ~ group + sex + sv1)

# diff analysis using deseq2
oligo_dds_diff = DESeq(oligo_dds, 
                       test = "Wald", 
                       fitType = "local",
                      # reduced = ~ sv1, # lrt only
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
oligo_dds_norm_counts = counts(oligo_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
oligo_dds_res = results(oligo_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
# resultsNames(oligo_dds_diff)

# # fold change shrinkage
# oligo_dds_res_shr <- lfcShrink(oligo_dds_diff, 
#                                coef = "group_pd_vs_control", 
#                                type = "apeglm")

# make results df
oligo_dds_res_df <- data.frame(chr = oligo_peak_coords_clean$chr,
                               start = oligo_peak_coords_clean$start,
                               end = oligo_peak_coords_clean$end,
                               baseMean = oligo_dds_res$baseMean,
                               lfc = oligo_dds_res$log2FoldChange,
                               lfcSE = oligo_dds_res$lfcSE,
                              # lfc_shr = oligo_dds_res_shr$log2FoldChange,
                               stat = oligo_dds_res$stat,
                               pvalue = oligo_dds_res$pvalue,
                               padj = oligo_dds_res$padj) 

# get only sig hyper acetylated peaks
oligo_dds_res_df_hyper <- oligo_dds_res_df %>%
  subset(padj < 0.05 & lfc > 1.5)

dim(oligo_dds_res_df_hyper)

# get only sig hypoacetylated peaks
oligo_dds_res_df_hypo <- oligo_dds_res_df %>%
  subset(padj < 0.05 & lfc < -1.5)

nrow(oligo_dds_res_df_hypo)
```


```{r oligo_qqplot}
qq_p_gg(oligo_dds_res_df$pvalue)
lambda_from_p(oligo_dds_res_df$pvalue)
```

```{r oligo_volcano}
ggplot(oligo_dds_res_df, 
       aes(x = lfc, 
           y = -log(pvalue))) +
  geom_point() +
  theme_bw() 
```

```{r oligo_go_terms_hyper}
# convert peaks to granges obj
oligo_dds_res_df_hyper_gr <- GRanges(seqnames = oligo_dds_res_df_hyper$chr,
                                     ranges   = IRanges(as.numeric(oligo_dds_res_df_hyper$start), 
                                                        as.numeric(oligo_dds_res_df_hyper$end)),
                                     strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(oligo_dds_res_df_hyper_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
oligo_dds_res_df_hyper_anno <- annotatePeak(oligo_dds_res_df_hyper_gr, 
                                            TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                            tssRegion=c(-3000,3000), 
                                            annoDb="org.Hs.eg.db")

# merge annotate to res df
oligo_dds_res_df_hyper <- merge(oligo_dds_res_df_hyper,
                                as.data.frame(oligo_dds_res_df_hyper_anno@anno) %>%
                                  mutate(chr = str_sub(seqnames, 4, 6)),
                                by = c("chr", "start", "end"))

oligo_hyper_genes <- unique(na.omit(as.data.frame(oligo_dds_res_df_hyper_anno)$ENSEMBL))

oligo_hyper_go <- enrichGO(gene=oligo_hyper_genes, 
                           OrgDb=org.Hs.eg.db, 
                           keyType="ENSEMBL",
                           ont="BP", 
                           #pvalueCutoff=0.01, 
                           qvalueCutoff=0.05)
dotplot(oligo_hyper_go)
```

```{r oligo_go_terms_hypo}
# convert peaks to granges obj
oligo_dds_res_df_hypo_gr <- GRanges(seqnames = oligo_dds_res_df_hypo$chr,
                                    ranges   = IRanges(as.numeric(oligo_dds_res_df_hypo$start), 
                                                       as.numeric(oligo_dds_res_df_hypo$end)),
                                    strand   = "*")

# harmonize chromosome naming style with TxDb (e.g., "chr1" vs "1")
seqlevelsStyle(oligo_dds_res_df_hypo_gr) <- seqlevelsStyle(TxDb.Hsapiens.UCSC.hg38.knownGene)

# annotate peaks
oligo_dds_res_df_hypo_anno <- annotatePeak(oligo_dds_res_df_hypo_gr, 
                                           TxDb=TxDb.Hsapiens.UCSC.hg38.knownGene, 
                                           tssRegion=c(-3000,3000), 
                                           annoDb="org.Hs.eg.db")

# merge annotate to res df
oligo_dds_res_df_hypo <- merge(oligo_dds_res_df_hypo,
                               as.data.frame(oligo_dds_res_df_hypo_anno@anno) %>%
                                 mutate(chr = str_sub(seqnames, 4, 6)),
                               by = c("chr", "start", "end"))

oligo_hypo_genes <- unique(na.omit(as.data.frame(oligo_dds_res_df_hypo_anno)$ENSEMBL))

oligo_hypo_go <- enrichGO(gene=oligo_hypo_genes, 
                          OrgDb=org.Hs.eg.db, 
                          keyType="ENSEMBL",
                          ont="BP", 
                          #pvalueCutoff=0.01, 
                          qvalueCutoff=0.05)
dotplot(oligo_hypo_go)
```


### Differential analysis of neuro

```{r neuro_diff_filter}
# remove sex chromosomes
neuro_feat_clean_noxy <- neuro_feat_clean %>%
  subset(!chr %in% c("X", "Y"))

# separate coordinates and counts
neuro_peak_coords <- neuro_feat_clean[, 1:3]
neuro_counts <- as.matrix(neuro_feat_clean[, -(1:3)])

# recode group as factor
# set control as base level
neuro_meta_svs <- neuro_meta_svs %>%
  mutate(group = relevel(factor(group), 
                         ref = "control")) 

# create edgeR dgelist for filtering
neuro_dgelist <- DGEList(counts = neuro_counts, 
                         samples = neuro_meta_svs)

# replicate-aware filtering - min count 5 
neuro_keep <- filterByExpr(neuro_dgelist, 
                           group = neuro_dgelist$samples$group, 
                           min.count = 5,
                           min.total.count=50) # smallest group (control) has 4 samples

# remove those with <5 from count and peaks
neuro_peak_coords_clean <- neuro_peak_coords[neuro_keep, ]
neuro_counts_clean <- neuro_counts[neuro_keep, ]

# recreate neuro dgelist
neuro_dgelist <- DGEList(counts = neuro_counts_clean, 
                         genes = neuro_peak_coords_clean,
                         samples = neuro_meta_svs)
```


```{r neuro_diff_deseq2}
# make deseq2 object
neuro_dds = DESeqDataSetFromMatrix(countData = neuro_counts_clean,
                                   colData = neuro_meta_svs,
                                   design = ~ group + sv1 + sv2)

# diff analysis using deseq2
neuro_dds_diff = DESeq(neuro_dds, 
                       test = "Wald", 
                       fitType = "local",
                      # reduced = ~ sv1, # lrt only
                       sfType = "poscounts")

# get normalized peak counts (uses method from DESeq)
neuro_dds_norm_counts = counts(neuro_dds_diff, 
                               normalized = TRUE)  

# create results dataframe
neuro_dds_res = results(neuro_dds_diff, 
                        pAdjustMethod = "BH",
                        independentFiltering = FALSE, 
                        altHypothesis = "greaterAbs")

# get result names
# resultsNames(neuro_dds_diff)

# # fold change shrinkage
# neuro_dds_res_shr <- lfcShrink(neuro_dds_diff, 
#                                coef = "group_pd_vs_control", 
#                                type = "apeglm")

# make results df
neuro_dds_res_df <- data.frame(chr = neuro_peak_coords_clean$chr,
                               start = neuro_peak_coords_clean$start,
                               end = neuro_peak_coords_clean$end,
                               baseMean = neuro_dds_res$baseMean,
                               lfc = neuro_dds_res$log2FoldChange,
                               lfcSE = neuro_dds_res$lfcSE,
                              # lfc_shr = neuro_dds_res_shr$log2FoldChange,
                               stat = neuro_dds_res$stat,
                               pvalue = neuro_dds_res$pvalue,
                               padj = neuro_dds_res$padj) 

```


```{r neuro_qqplot}
qq_p_gg(neuro_dds_res_df$pvalue)
lambda_from_p(neuro_dds_res_df$pvalue)
```

```{r neuro_volcano}
ggplot(neuro_dds_res_df, 
       aes(x = lfc, 
           y = -log(pvalue))) +
  geom_point() +
  theme_bw() 
```